Determinism, Complexity, and Predictability in Computer Performance

### 2014-03-17 methods.tex
modified methods.tex to get the '~21,000 cycles' stuff right

\svd is a Fortran program from the LAPACK linear algebra package
\cite{lapack}.  It calculates the singular value decomposition of an
rectangular $M$ by $N$ real-valued matrix.  For our experiments, we
choose $M=750$ and $N=1000$ and generated the matrix entries randomly.
% 
% \footnote{Multiple randomly generated matrices were
%   investigated but no measurable effect was present in the resulting
%   time series.}. 
% 
The behavior of this program as it computes the singular values of
this matrix is very interesting, as shown in Figure
\ref{fig:svd-ts-colored}.  As the code moves though its different
phases---diagonalizing the matrix, computing its transpose,
multiplying, etc.---the processor utilization changes quite radically,
as shown in Figure~\ref{fig:svd-ts-colored}.
\begin{figure}[t]
    \centering
    \includegraphics[width=\columnwidth]{figs/SVD1RegimesColored}
    \caption{The instructions executed per CPU clock cycle (IPC)
      during the execution of \svd. Each point is the average IPC
      during a 100,000 instruction period.  The colors refer to different
``regimes'' of the signal, as discussed in the text.}
    \label{fig:svd-ts-colored}
  \end{figure}
For the first $\sim$27,000 instructions{\color{red}[[Joshua,I am not sure about this, I think it is more instructions than that. 27,000 instructions are aggregated  in the first point of the time series (every 100,000 instructions is 1 point)...Liz how did you calculate this, I think it should be 21,000*100,000]]}, roughly 1.8 instructions are
executed per cycle, on the average, by the eight processing units on
this chip.  After that, the IPC moves through a number of different
oscillatory regimes, which we have color-coded in the figure in order
to make textual cross-references easy to track and understand.

>>>>

\svd is a Fortran program from the LAPACK linear algebra package
\cite{lapack}.  It calculates the singular value decomposition of an
rectangular $M$ by $N$ real-valued matrix.  For our experiments, we
choose $M=750$ and $N=1000$ and generated the matrix entries randomly.
% 
% \footnote{Multiple randomly generated matrices were
%   investigated but no measurable effect was present in the resulting
%   time series.}. 
% 
The behavior of this program as it computes the singular values of
this matrix is very interesting, as shown in Figure
\ref{fig:svd-ts-colored}.  As the code moves though its different
phases---diagonalizing the matrix, computing its transpose,
multiplying, etc.---the processor utilization changes quite radically,
as shown in Figure~\ref{fig:svd-ts-colored}.
\begin{figure}[t]
    \centering
    \includegraphics[width=\columnwidth]{figs/SVD1RegimesColored}
    \caption{The instructions executed per CPU clock cycle (IPC)
      during the execution of \svd. Each point is the average IPC
      during a 100,000 instruction period.  The colors refer to different
``regimes'' of the signal, as discussed in the text.}
    \label{fig:svd-ts-colored}
  \end{figure}
For the first $\sim$21,000 measurements (21,000 $\times$ 100,000
cycles), roughly 1.8 instructions are executed per cycle, on the
average, by the eight processing units on this chip.  After that, the
IPC moves through a number of different oscillatory regimes, which we
have color-coded in the figure in order to make textual
cross-references easy to track and understand.
